---
title: "Black_Friday_Project"
author: "Irene Huang and Kathleen Gendotti"
subtitle: MGSC 310, Fall 2018, Professor Hersh
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Changes the age, city category and occupation variables from integers to factors, provides the variable types and checks for any missing values, sets seed, gets rid of scientific notation
```{r}
BlackFridayDF <- read.csv("/Users/kathleengendotti/Downloads/BlackFriday.csv")
BlackFridayDF$Age <- as.factor(BlackFridayDF$Age)
BlackFridayDF$City_Category <- as.factor(BlackFridayDF$City_Category)
BlackFridayDF$Occupation <- as.factor(BlackFridayDF$Occupation)
colSums(is.na(BlackFridayDF))
BlackFridayDF <- BlackFridayDF[!is.na(BlackFridayDF$Product_Category_1), ]
BlackFridayDF <- BlackFridayDF[!is.na(BlackFridayDF$Product_Category_2), ]
BlackFridayDF <- BlackFridayDF[!is.na(BlackFridayDF$Product_Category_3), ]
BlackFridayDF <- BlackFridayDF[!is.na(BlackFridayDF$User_ID), ]
BlackFridayDF <- BlackFridayDF[!is.na(BlackFridayDF$Product_ID), ]
colSums(is.na(BlackFridayDF))
set.seed(1861)
options(scipen=15)
```
Gives mean, median, 25th and 75th quartiles, min, and max of the variables
```{r}
summary(BlackFridayDF)
```

Histogram of purchase
```{r}
hist(BlackFridayDF$Purchase, col = 8, xlab = "Purchase", ylab = "Count", main = "Purchase vs. count")
```


This pie chart shows the distribution of total purchases by age group
```{r}
library('plotly')
pieAge <- plot_ly(BlackFridayDF, labels = ~Age, values = ~Purchase, type = 'pie') %>%
  layout(title = 'Age Group Purchases',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pieAge
```

Pie chart of purchases by gender
```{r}

pieGender <- plot_ly(BlackFridayDF, labels = ~Gender, values = ~Purchase, type = 'pie') %>%
  layout(title = 'Gender Group Purchases',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
pieGender
```

These boxplots shows the distribution of purchases by gender
```{r}
ggplot(aes(x=Gender, y=Purchase, fill=Gender), data = BlackFridayDF)+geom_boxplot()
```

Histogram plots of purchase by occupation
```{r}
BlackFridayDF$Purchase <- log(BlackFridayDF$Purchase)
library("ggplot2")
ggplot(aes(Purchase), data = BlackFridayDF) + geom_histogram() + theme_light() + labs(title = "Plot of Total Purchases by Occupation", x = "Purchase (in dollars)") + facet_wrap(~Occupation)
```

This plot shows the frequency of various total purchase values colored by the number of years the individual has lived in their current city.
```{r}
ggplot(aes(Purchase, fill = City_Category), data = BlackFridayDF) + geom_histogram() + theme_light() + labs(title = "Plot of Total Purchase Price Frequency by Number of Years Lived in Current City", x = "Purchase (in dollars)")+ facet_wrap(~Stay_In_Current_City_Years)
```


These plots show the plots of purchases separated by the city catgeories and colored in by the number of years the individual has lived in their current city.
```{r}
ggplot(aes(Purchase, fill = Stay_In_Current_City_Years), data = BlackFridayDF) + geom_histogram() + theme_light() + labs(title = "Plot of Total Purchase Price Frequency by City Catgeory", subtitle = "Colored by Number of Years Lived in Current City", x = "Purchase (in dollars)") + facet_wrap(~City_Category)

```


Linear Regression Model
```{r}
Linear_regression <- lm(formula = Purchase ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status, data = BlackFridayDF)
summary(Linear_regression)
```

Generalized Linear Model
```{r}
library(glmnet)
logitFit <- glm(Purchase ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status, data = BlackFridayDF)
summary(logitFit)

BlackFridayDF$PurchaseHigh <- ifelse(BlackFridayDF$Purchase > median(BlackFridayDF$Purchase),1,0)

doBy::summaryBy(PurchaseHigh ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status, data = BlackFridayDF)

tFit <- glm(PurchaseHigh ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status, data = BlackFridayDF, family = binomial)
summary(tFit)
```

Lasso vs. Ridge Models
```{r}
myFormula <- as.formula(Purchase ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status)

library("useful")
Xvar <- build.x(myFormula, data=BlackFridayDF)
Yvar <- build.y(myFormula, data=BlackFridayDF)
library("glmnet")
LassoMod <- cv.glmnet(x = Xvar, y = Yvar, alpha = 1, nfolds=10)
RidgeMod <- cv.glmnet(x = Xvar, y = Yvar, alpha = 0, nfolds=10)

Lassomin = coef(LassoMod, s = "lambda.min")
Lasso1se = coef(LassoMod, s = "lambda.1se")
Ridgemin = coef(RidgeMod, s = "lambda.min")
Ridge1se = coef(RidgeMod, s = "lambda.1se")

Lassomin = round(Lassomin, digits = 4)
Ridgemin = round(Ridgemin, digits = 4)
Lasso1se = round(Lasso1se, digits = 4)
Ridge1se = round(Ridge1se, digits = 4)

table1 = cbind(Lassomin,Ridgemin)
colnames(table1) = c("Lasso", "Ridge")
print(table1)

table2 = cbind(Lasso1se, Ridge1se)
colnames(table2) = c("Lasso", "Ridge")
print(table2)

r2Lassomin <- LassoMod$glmnet.fit$dev.ratio[which(LassoMod$glmnet.fit$lambda == LassoMod$lambda.min)]
r2Lasso1se <- LassoMod$glmnet.fit$dev.ratio[which(LassoMod$glmnet.fit$lambda == LassoMod$lambda.1se)]
r2Lassomin
r2Lasso1se

r2Ridgemin <- RidgeMod$glmnet.fit$dev.ratio[which(RidgeMod$glmnet.fit$lambda == RidgeMod$lambda.min)]
r2Ridge1se <- RidgeMod$glmnet.fit$dev.ratio[which(RidgeMod$glmnet.fit$lambda == RidgeMod$lambda.1se)]
r2Ridgemin
r2Ridge1se
```

Forward Stepwise Model
```{r}
library("leaps")
#was not sure on good value to set for nvmax so went with the default of 8
BFFitFwd <- regsubsets(Purchase ~ Gender + Age + Occupation + City_Category + Stay_In_Current_City_Years + Marital_Status , data = BlackFridayDF, nvmax = 8, method = "forward")
summary(BFFitFwd)

#created separate variables for city catrgories B and C since the forward stepwise model chose them
BlackFridayDF$City_CategoryB <- ifelse(BlackFridayDF$City_Category == "B",1,0)
BlackFridayDF$City_CategoryC <- ifelse(BlackFridayDF$City_Category == "C",1,0)
#created separate variable for the age group of people between 51 and 55
BlackFridayDF$Age51 <- ifelse(BlackFridayDF$Age == "51-55",1,0)
#created separate variables for the occupations the model chose
BlackFridayDF$Occupation1 <- ifelse(BlackFridayDF$Occupation == "1",1,0)
BlackFridayDF$Occupation10 <- ifelse(BlackFridayDF$Occupation == "10",1,0)
BlackFridayDF$Occupation19 <- ifelse(BlackFridayDF$Occupation == "19",1,0)
BlackFridayDF$Occupation20 <- ifelse(BlackFridayDF$Occupation == "20",1,0)

lmDayDF <- lm(Purchase ~ Gender + City_CategoryB + City_CategoryC + Age51 + Occupation1 + Occupation10 + Occupation19 + Occupation20, data = BlackFridayDF)
summary(lmDayDF)
```


Calculates RMSE and MSE over all the models
```{r}
RMSE = function(x1,x2) {
  sqrt(mean(x1-x2)^2)
}

lmRMSE = RMSE(predict(Linear_regression, newx = Xvar), BlackFridayDF$Purchase)
ltRMSE = RMSE(predict(logitFit, newx = Xvar), BlackFridayDF$Purchase)
LassoRMSE = RMSE(predict(LassoMod, newx = Xvar), BlackFridayDF$Purchase)
RidgeRMSE = RMSE(predict(RidgeMod, newx = Xvar), BlackFridayDF$Purchase)
lmDayDFRMSE = RMSE(predict(lmDayDF, newx = Xvar), BlackFridayDF$Purchase)

MSE <- function(yhat, ytrue){ 
  mean((yhat - ytrue)^2)
}

lmMSE = MSE(predict(Linear_regression, newx = Xvar), BlackFridayDF$Purchase)
ltMSE = MSE(predict(logitFit, newx = Xvar), BlackFridayDF$Purchase)
LassoMSE = MSE(predict(LassoMod, newx = Xvar), BlackFridayDF$Purchase)
RidgeMSE = MSE(predict(RidgeMod, newx = Xvar), BlackFridayDF$Purchase)
lmDayDFMSE = RMSE(predict(lmDayDF, newx = Xvar), BlackFridayDF$Purchase)

compareModels <- matrix(c(lmRMSE,ltRMSE, LassoRMSE,RidgeRMSE, lmDayDFRMSE, lmMSE, ltMSE, LassoMSE,RidgeMSE, lmDayDFMSE),ncol=2,byrow=TRUE)
colnames(compareModels) <- c("RMSE","MSE")
rownames(compareModels) <- c("Linear Regression", "Generalized Linear","Lasso","Ridge", "Forward Stepwise")
compareModels <- as.table(compareModels)
compareModels
```



